---
title: "Midterm Exam"
author: "Jingwen Xu"
date: "11/2/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(magrittr)
library(rstanarm)
library(pwr)
knitr::opts_chunk$set(echo = TRUE)
```

## Instruction

This is your midterm exam that you are expected to work on it alone.  You may NOT  discuss any of the content of your exam with anyone except your instructor. This includes text, chat, email and other online forums.  We expect you to respect and follow the [GRS Academic and Professional Conduct Code](http://www.bu.edu/cas/files/2017/02/GRS-Academic-Conduct-Code-Final.pdf). 

Although you may NOT ask anyone directly, you are allowed to use external resources such as R codes on the Internet.  If you do use someone's code, please make sure you clearly cite the origin of the code.

When you finish, please compile and submit the PDF file and the link to the GitHub repository that contains the entire analysis.


## Introduction

In this exam, you will act as both the client and the consultant for the data that you collected in the data collection exercise (20pts).  Please note that you are not allowed to change the data.  The goal of this exam is to demonstrate your ability to perform the statistical analysis that you learned in this class so far.  It is important to note that significance of the analysis is not the main goal of this exam but the focus is on the appropriateness of your approaches.

### Data Description (10pts)

Please explain what your data is about and what the comparison of interest is.  In the process, please make sure to demonstrate that you can load your data properly into R.

```{r }
# The introduction of my data
## hair_loss: When I took the survey, I set a 6 points scale to assess the severity of hair loss of respondents. To avoid categorical outcomes, I assume “hair_loss<=3” to be indicator “0” which means that the respondents has little concern about hair loss and “hair_loss>3” to be indicator “1”.
## gender: “1” indicates female, “0” indicates male.
## age: “1” indicates 18-25 years old, “2” indicates 26-40 years old, “3” indicates more than 40 years old.
## insomnia: “1” indicates that respondents have insomnia, “0” indicates no insomnia.
## sleep_t: The average daily sleeping time. “1” indicates less than 4 hours, “2” indicates 4-6 hours, “3” indicates 6-8 hours, “4” indicates more than 8 hours.
## computer_t: The average daily computer facing time. “1” indicates less than 1 hour, “2” indicates 1-3 hours, “3” indicates 3-5 hours, “4” indicates 5-8 hours, “5” indicates more than 8 hours.
## sport_t: The average daily sport time. “1” indicates less than 1 hour, “2” indicates 1-2 hours, “3” indicates 2-4 hours, “4” indicates more than 4 hours.
## genetic: “1” indicates that the respondent has family hereditary hair loss, “0” indicates none.
## pregnant: “1” indicates that the respondent is in the post-pregnancy stage, “0” indicates none.
## menopause: “1” indicates that the respondent is menopause, “0” indicates none.
## chemical: “1” indicates that the respondent has used poor quality hair dye or perm, “0” indicates none.
## disease: “1” indicates that the respondent has hair follicle disease, “0” indicates none.


# The comparison of interest
## Given the information of different respondents such as their gender, age, having insomnia or not and so on, compare their probability to have concern of hair loss.

# Load and read my data into R
hairloss <- read.csv("https://raw.githubusercontent.com/VivianXu66/Midterm_Exam/main/collection.csv",header=T)
head(hairloss)
```

### EDA (10pts)

Please create one (maybe two) figure(s) that highlights the contrast of interest.  Make sure you think ahead and match your figure with the analysis.  For example, if your model requires you to take a log, make sure you take log in the figure as well.

```{r }
summary(hairloss)

# Simply organize the data
## According to common sense, the predictor "genetic", "pregnant", "menopause", "chemical" and "disease" are all special contributes to hair loss which can disturb the effect of other normal contributes on the outcome.
hairloss <- subset(hairloss,hairloss$genetic=="0"&hairloss$pregnant=="0"&hairloss$menopause=="0"&hairloss$chemical=="0"&hairloss$disease=="0")
hairloss %<>% select(-c(1,9:13))
rownames(hairloss) <- c(1:51)


# There are many predictors in this data. I want to operate EDA on the predictor "computer_t" and the outcome of interest as an example.
## Firstly, get the counts of hair loss or not in different computer_t groups.
bar_data <- hairloss %>% group_by(hair_loss,computer_t) %>% summarise(Count=n())

## Plot the stacked bar
computer_time <- c(rep("less than 1 hour",2),rep("1-3 hours",2),rep("3-5 hours",2),rep("5-8 hours",2),rep("more than 8 hours",2))
loss_or_not <- rep(c("hair loss","not hair loss"),5)
value <- c(3,4,0,10,3,8,2,11,3,7)
data <- data.frame(computer_time,loss_or_not,value)
ggplot(data,aes(fill=loss_or_not,y=value,x=computer_time)) + 
    geom_bar(position="stack",stat="identity")+
  theme(axis.text.x=element_text(angle=30,hjust=1),
        axis.text=element_text(size=11),
        axis.title=element_text(size=13,face="bold"))+ggtitle("The counts of hair loss or not for each computer time group")+xlab("average daily computer facing time")+ylab("Hair loss or not")

## Plot the percentage stacked bar
ggplot(data,aes(fill=loss_or_not,y=value,x=computer_time)) + 
    geom_bar(position="fill",stat="identity")+
  theme(axis.text.x=element_text(angle=30,hjust=1),
        axis.text=element_text(size=11),
        axis.title=element_text(size=13,face="bold"))+ggtitle("The percentage of hair loss or not for each computer time group")+xlab("average daily computer facing time")+ylab("Hair loss or not")
```

### Power Analysis (10pts)

Please perform power analysis on the project.  Use 80% power, the sample size you used and infer the level of effect size you will be able to detect.  Discuss whether your sample size was enough for the problem at hand.  Please note that method of power analysis should match the analysis.  Also, please clearly state why you should NOT use the effect size from the fitted model.

```{r }
# Use the function in pwr package to perform power analysis
pwr.p.test(h=NULL,n=51,sig.level=0.05,power=0.8)
## The effect size h = 0.3923029


# Cohen suggests that h values of 0.2, 0.5, and 0.8 represent small, medium, and large effect sizes respectively.(From:https://www.statmethods.net/stats/power.html)
par(mar=c(3,3,2,1), mgp=c(2,.7,0), tck=-.01)
nlist<-2:100
powres<-sapply(nlist,function(x)pwr.p.test(n=x,h=0.5,sig.level=0.05,NULL)$power)
plot(nlist,powres,type="l",xlab="sample size",ylab="power");abline(h=0.8,lty=2); abline(v=nlist[which(powres>0.8 )[1]])## The original code is from https://bu-rstudio-connect.bu.edu/connect/#/apps/15/access
## From the plot, we can see that if effect size is medium and power is above 0.8, we need to ensure the sample size above 30. In my data, sample size equals to 51 which is enough for the problem.


# Effect size is usually hypothesized due to the unknown true value.

```


### Modeling (10pts)

Please pick a regression model that best fits your data and fit your model.  Please make sure you describe why you decide to choose the model. Also, if you are using GLM, make sure you explain your choice of link function as well.
```{r }
fit <- stan_glmer(hair_loss~gender+insomnia+(1|age)+(1|sleep_t)+(1|computer_t)+(1|sport_t),data=hairloss,family=binomial,refresh=0)
fit
ranef(fit)
## This is a multilevel logistic model. Firstly, the outcome in the data is binary so we need to fit a logistic model. Then considering that there are several index variables in the predictors, I chose the multilevel model because the classical regression is no pooling which may overstate the difference between the groups.
```

### Validation (10pts)

Please perform a necessary validation and argue why your choice of the model is appropriate.  

```{r }
# Leave-one-out cross validation
logis_loo <- loo(fit)
logis_loo
logis_loo_1 <- loo(fit,k_threshold = 0.7)
logis_loo_1
## There is one unstable observation for the first loo. But after refitting the model 1 times, we can see that all pareto k estimates are ok. The estimated log score is -26.0 which means the fit is pretty well. The effective number of parameters is 8.1 and there are just 8 parameters in this data.

# To better excludes the unstable observation in LOO, I use the k-fold cross validation
kfold(fit,K=10)
## The results are similar to that of LOO valiation.

# plot the posterior predictive checks
pp_check(fit)
## According to the plot, the model captures certain characteristics of the data so that they have the similar curve trend despite of the deviance.
```


### Inference (10pts)

Based on the result so far please perform statistical inference to compare the comparison of interest.

```{r }
# Summary the inference
summary(fit)
## We can see that except for intercept and the coefficient of gender, other coefficients are not significant.

# Maybe a prediction simulation is better for inference, but I don't know how to do such a simulation for multilevel logistic regression.
```


### Discussion (10pts)

Please clearly state your conclusion and the implication of the result.


1.Conclusion: (1)According to the coefficient, the people who have insomnia will be more likely to be bothered by hair loss. It's reasonable referring to the common sense.
(2)According to the coefficient, the probability that women concern about hair loss is larger than men do.
(3)From the EDA analysis and model coefficient, we can see that there are relationship between computer_t(sport_t, sleep_t and age) and hair loss. But we need larger sample to determine the precise positive or negative relationship.

2.Implication: In China, hair loss problem has became a popular topic especially among the young people due to their increasing life pressure. Either job or study requires a lot of time facing computer which may results in hair loss. And sedentariness can reduce metabolism. Besides, more and more people suffer from insomnia. I want to clarify the relationship between these phenomenons and hair loss to alert everyone and promote a healthy lifestyle.

### Limitations and future opportunity. (10pts)

Please list concerns about your analysis.  Also, please state how you might go about fixing the problem in your future study.


1.Concerns: (1)The data is not large enough with only 56 observations so that it's not persuasive with certain missing value in certain groups. For example, There is only one respondent from group 2 of computer_t having concern about hair loss. But her hair loss is mostly related to the poor quality hair dye or perm. After removing such a disturb, there is no respondent from group 2 of computer_t having concern about hair loss. It's not reasonable and persuasive.  
(2)The survey is mainly issued in my Wechat where most of the people are 18-25 or more than 40 years old so the sample is not random enough.
(3)The judge of hair loss or not is based on the subjective consciousness of the respondents which may lead to measurement error.
(4)The LOO cross validation may be not proper for multilevel model. There is a validation method called leave-one-group-out which will be better for multilevel model.

2.Possible fixing solutions:(1)Take the survey in more general platform which can ensure the randomness. And find more respondents to do the survey.
(2)Clarify the evaluation about hair loss or not with more objective and scientific criterion. For example, I can ask the professional doctor for help to set the criterion.
(3)There may be interaction among computer_t, sleep_t and sport_t, I should study if the interaction exists or not for fitting model more precisely.
(4)I think I should simplify the model in further study.


### Comments or questions
If you have any comments or questions, please write them here.


Question: (1)Is there a good way to deal with the overlapping of factor baseline and intercept if the predictors have more than one index variables?
(2)How to properly choose the function for power analysis?

